### 2023-03-16 chapters/paper_review/03_approaches.tex
Change reference keys for review paper (WIP)

\section{Speech Representation Learning Paradigms} \label{sec:approach}
% {\color{blue} Reviewer: Hung-yi, Daniel , Abdo}\\
% {\color{blue} Hung-yi + Abdo}\\
%Start by discussing central challenges for self-supervised representation learning in speech: 1)Can't do instance classification in variable-length sequences. 2)No pre-defined vocabulary. 3)No pre-defined segmentation. \\
%New approaches offer deep contextualization of learned representations. \\
% {\color{red} first version for a list of symbols to use -- let's keep updating it as needed -- Abdo}\\
Due to the characteristics of speech, SSL pretext tasks developed for CV
and NLP may not directly apply to speech.
Below we summarize the characteristics of speech as compared to CV and NLP.
\begin{itemize}
\item \textit{Speech is a sequence.} 
Unlike CV, in which an image usually has a fixed size representation, it is
natural to represent a speech utterance as a variable-length sequence. 
Therefore, pretext tasks developed for CV cannot generally be directly
applied to speech.

\item \textit{Speech is a long sequence without segment boundaries.} 
Both text and speech can be represented as sequences. From this viewpoint, it
is natural to apply learning approaches developed for text directly to
speech. 
% However, the sequence representations of speech are usually much longer than those of text and they contain no natural boundaries at which they can be split into segments.
In NLP, morpheme-like tokens are widely used as sequence units in
pre-training. The standard BERT takes 512 morpheme-like tokens as
input, usually covering a paragraph including several sentences. 
% In speech processing, the audio signal is recorded as a waveform represents the amplitude of pressure. 
However, speech signals consist of sound pressure measurements with thousands
of samples per second, resulting in sequences much longer than those for text. Even
spectral representations which reduce the sequence length can have hundreds of
frames per second.
%This raw signal is often converted to a spectral representation where each frame represents 20-25 milliseconds of audio computed with overlaps at 10 millisecond strides.
% a frame is the typical sequence unit in speech, and each frame usually represents a length of 10 milliseconds. 
% Therefore, an utterance is usually represented as a long sequence with more than a thousand frames. 
Processing such sequences with typical neural network architectures like
Transformers can result in problems with running time and memory requirements. 
One could gather consecutive frames to form shorter segments,
but unlike text, there is no obvious segmentation for unlabeled
speech.
\item \textit{Speech is continuous.} 
In NLP, it is common to use a pretext task that models a categorical
distribution of masked or future inputs. Since text is easily broken down into
individual tokens such as words, subwords, or characters, it is
straightforward to define a finite vocabulary for such tasks.
%The generative-based pretexts are widely applied in NLP.
%Because a sentence is composed of tokens, sentence generation can be formulated as a series of classification tasks.  
However, this idea does not apply to speech modeling because speech
signals are continuous; 
% thus, such a vocabulary is unavailable.
  in this sense there is no such thing as a speech vocabulary.   % AMH: check
\item \textit{Speech processing tasks are diverse.}
Building generalizable self-supervised representation models for diverse speech
processing tasks is challenging.
Speech contains rich, hierarchical information, and different speech tasks
may require mutually orthogonal information.
For example, speech recognition requires a model that extracts content information
but ignores speaker information; in contrast, speaker recognition 
requires a model that extracts speaker information but removes content information.
Therefore, it is challenging to define a self-supervised model whose
representations are suitable for both speech recognition and speaker
recognition. Analogous considerations apply within CV and NLP.
\end{itemize}
%(Although we list the problems here, we do not highlight the solutions in the following paragraph.)

In the sections below, we group modern SSL pretext tasks designed for speech
into three main categories: \textit{generative} approaches,
\textit{contrastive} approaches and \textit{predictive} approaches. 
%For each category, we provide a brief motivation and then describe prominent models that belong to it. Lastly, we discuss a set of challenges related to the category.
\Cref{fig:timeline} shows a timeline of the models covered in these sections
with each model colored according to our categorization. 
\Cref{table:pretext} summarizes model pretext tasks along within the categories.

%\input{tables/model}

% \begin{center}
% \begin{tabular}{ |c|c| } 
%  \hline
%  $s$ & a single waveform sample \\
%  $S$ & a sequence of waveform samples \\
%  $x$ & a single feature vector (MFCC or learned)\\
%  $X$ & a sequence of feature vectors \\
%  $\widetilde{X}$ & a data augmented sequence of feature vectors \\
%  $z$ & a single latent vector \\
%  $Z$ & a sequence of latent vectors \\
%  $h$ & a single hidden vector representation \\
%  $H$ & a sequence of hidden vector representation \\
%  $m$ & a binary mask for a single time step \\
%  $M$ & a binary mask for the whole sequence \\
%  $y$ & a single supervised label \\
%  $Y$ & a sequence of supervised labels \\
%  $\tau$ & temperature variable \\ 
%  $f()$ & an encoding network - use $f_1()$ $f_2()$ if multiple\\
%  $g()$ & a decoding network \\
%  $P()$ & Probability \\
%  $E()$ & Expectation \\
%  $L()$ & Loss \\
%  $\mathcal{L}$ & Likelihood \\
%  \hline
% \end{tabular}
% \end{center}

@@

%A few other conventions should be noted here. 
We use $Q = \{q_1,q_2, ..., q_T\}$ to denote representations that are quantized via codebook learning. Alternatively, discrete representations may take the form of one-hot vectors, or the equivalent integer IDs, which we denote by $C = \{c_1,c_2, ..., c_T\}$. We use a circumflex to denote that, for instance, $\hat{x}_t$ is an approximation of $x_t$. Finally, we often use a subscript when defining a loss, $\mathcal{L}_i$, to imply that the total loss is computed as a sum over $i$, unless otherwise stated.

@@

\edit{After pre-training,} the representation model $f(\cdot)$ can be fine-tuned for a downstream task directly or used to extract features which are fed to another model, as visualized in \cref{fig:SSL_framework}. It is not uncommon to use the output representation~$H$, but often representations from hidden layers of $f(\cdot)$ are better suited \cite{pasad2021}.

\begin{figure*}[t!]
    \centering
    \includegraphics[width=0.98\textwidth]{paper_review/model_timeline.pdf}
   % \includegraphics[width=0.98\textwidth]{paper_review/model_timeline_wo_dccae.pdf}
	 \caption{A selection of models listed according to first publication date
	 on arXiv or conference submission date when this clearly precedes the
	 former. The models are categorized as generative, contrastive, or predictive.
	 In addition, some models are characterized as embedding models or
	 multi-modal models, although most learn frame-level
	 representations from speech only. Some models use a mixture of generative
	 and contrastive tasks. For instance, PASE and PASE+ use a multi-task setup,
	 but find that generative tasks are the most important for downstream
	 task performance~\cite{pascual2019learning}.}
    \label{fig:timeline}
\end{figure*}

@@

% {\color{blue} Hung-yi}\\

@@

%In this category, the pretext task is to generate speech signals given quantized, past, or unmasked time steps. To solve the generative-based pretext, the model leverages an encoder $f(\cdot)$ and a decoder $g(\cdot)$. Given waveform samples $S$ or its corresponding acoustic feature vectors $X$, the encoder $f(\cdot)$ outputs hidden representations $H$. The waveform samples $S$ or acoustic feature vectors $X$ are usually corrupted before entering the encoder $f(\cdot)$. In this subsection, we only consider the acoustic features $X$ and ignore waveform samples $S$ for simplicity, but taking waveform samples as input and generating waveform samples is possible. Some regularization in the encoder $f(\cdot)$ or decoder $g(\cdot)$ is required to encourage the model to utilize global speech information and prevent naively copying input when reconstructing, especially when the input of $f(\cdot)$ is not corrupted. After pre-training, $f(\cdot)$ is used as the representation model, which takes a sequence of acoustic feature vectors $X=\{x_1,x_2, ..., x_T\}$ as input, and outputs hidden contextual representations $H =\{h_1,h_2,...,h_T\}$ for downstream tasks.

%In this category, the pretext task is to generate, or reconstruct, the input data based on some limited view. This includes predicting future inputs from past, masked from unmasked, or the original from some other corrupted view. Models that solve a generative pretext task typically consist of an encoder $f(\cdot)$ and a decoder $g(\cdot)$. Given an acoustic input $X =\{x_1,x_2, ..., x_T\}$, the encoder $f(\cdot)$ outputs a representation $H =\{h_1,h_2,...,h_T\}$. The decoder $g(\cdot)$ takes $H$ as input to generate parts of the original acoustic input. The input $X$ may be either the raw waveform samples or a sequence of spectral feature vectors. In the following, we will not distinguish between the two, but in practice, both are viable options. After pre-training, $f(\cdot)$ is used as the representation model. The model can be fine-tuned for a downstream task directly or used to extract features which are fed to another model, as visualized in \cref{fig:SSL_framework}. It is common to use the output representations $H$, but often representations from hidden layers of $f(\cdot)$ are more suited.

%All the generative approaches are summarized in \cref{table:generative}. 
%There is little work generating waveform samples $S$ because generating them is computationally expensive.
In this category, the pretext task is to generate, or reconstruct, the input data based on some limited view. This includes predicting future inputs from past inputs, masked from unmasked, or the original from some other corrupted view.  ``Generative'' as used in this paper hence refers to models that target the original input in their pretext task. Note that this differs from generative models, which learn \ci{distributions that allow} to sample new data.
% It is important to note that many of the approaches discussed in this subsection are ''reconstructive``, i.e., they focus on reconstructing the input signal rather than learning a probabilistic model for sampling novel sequences\footnote{Generative self-supervised models are not the same as generative probabilistic models that permit sampling new data.}.

%\footnote{The lengths of $X$ and $H$ are not necessary to be equal. Usually, $H$ is shorter than $X$. For simplicity of discussion, we assume the lengths of $X$ and $H$ are the same in this subsection.} % Any one care about this?

@@

%\subsubsection{Autoregressive reconstruction}
%\subsubsection{Autoregressive Prediction}
%\label{sssec:autoregressive}

\paragraph{Autoencoding} 
Since their introduction in the mid-1990s~\cite{hinton_94}, autoencoders (AEs)
have played an essential role in learning distributed latent representations of
sensory data. 
As described above, AEs consist of an encoder and decoder; the pretext task
is to reconstruct the given input. The most common type of AE places an
information bottleneck on the latent representation by simply having fewer
hidden units available than input features. This forces the model to discard
low-level details and disccourages the learning of trivial solutions. Other models add
regularization to the latent space to further improve the quality of the
learned representations.
% An AE converts an input utterance $X$ to a sequence of representations $H$ and then uses a decoder network to convert $H$ into an approximate reconstruction of the input $X$. Regularization is required to learn general and efficient latent representations and to avoid learning a trivial identity mapping. 
For instance, denoising autoencoders (DAEs) learn latent representations by
reconstructing from input corrupted by noise~\cite{DAE}. 
\edit{The Variational Autoencoder (VAE) is a probabilistic version of the AE which defines the latent representation via a posterior distribution over stochastic latent variables \cite{vae,rezende2014stochastic}. VAEs have been applied to speech in numerous works \cite{chung_recurrent_2015, fraccaro_sequential_2016, hsu2017learning, hsu2017unsupervised, aksan_stcn_2019}}.
The vector-quantized variational autoencoder (VQ-VAE) is another model in this
category~\cite{vqvae};
% The first method to sucessfully learn meaningful discrete latent representations was the Vector Quantised Variational Autoencoder (VQ-VAE)~\cite{vqvae}. 
it extends the original VAE~\cite{vae} with a novel
parameterization of the posterior distribution for discrete latent
representations. 
\edit{The VQ-VAE has been instrumental in generative speech modelling and recent work on generative spoken language modeling has successfully combined the idea of a discrete latent space with self-supervised learning \cite{SpeechResynthesis_IS21, pGSLM, GSLM}.}
% Forcing the use of discrete latent representations encourages the model to focus on salient input features, e.g., phonemes in speech, rather than using the capacity to model channel noise and imperceptible local input details.

Specifically, in the VQ-VAE, the continuous representation vector $h_t$ at the output of the encoder is quantized by mapping it to a codebook vector, which is then used as the input to the decoder. This operation is non-differentiable and the \edit{gradients of the loss with respect to the encoder parameters} must be obtained by approximation. In the VQ-VAE this is done using the straight-through estimator~\cite{bengio2013estimating}, i.e., the gradients \edit{with respect to the} encoder output are taken to be equal to those \edit{with respect to the} decoder input \ci{(i.e., the quatization step is ignored)}. Given a learned codebook $A\in\mathbb{R}^{K \times D}$, where $K$ is the codebook size and $D$ is the dimensionality of each codebook vector $a_k$, the quantized representation $q_t$ of $h_t$ is obtained as
\begin{align}
    q_t = a_k, \text{ where } k=\arg\min_j \norm{h_t - a_j}_2 .
\end{align}
\edit{The decoder $g(\cdot)$ is an autoregressive model that takes $q_{[1,t]}$ as input to generate $x_t$ \cite{oord2016wavenet}}. 
Codebook learning is facilitated by a two-term auxiliary loss similar to
classical vector quantization dictionary 
learning~\cite{burton_generalization_1983, soong_vector_1985}. 
Gradients for the codebook vectors are given solely by a term that moves
codebook vectors $a_k$ closer to the non-quantized vectors $h_t$. A so-called
\emph{commitment term} is added to ensure that non-quantized vectors do not grow
unboundedly by enforcing the encoder to keep them close to a codebook vector.
This commitment term is optimized only by the encoder. The
total VQ-VAE loss \edit{for a single timestep} is
% \begin{align}
%     %\mathcal{L} = \log p(x | q) +  \underset{\text{vq}}{\underbrace{\norm{\text{sg}\left[v\right] - A}_2^2}} + \underset{\text{commitment}}{\underbrace{\beta\norm{v - \text{sg}\left[A\right]}_2^2}} \enspace ,
%     \resizebox{.44\textwidth}{!}{%
%     $\mathcal{L}_t = \underset{\text{encoder+decoder}}{\underbrace{\log p(x_t | q_{[1,t]})}} +  \underset{\text{codebook}}{\underbrace{\norm{\mathrm{sg}\left[h_t\right] - A}_2^2}} + \beta\underset{\text{encoder}}{\underbrace{\norm{h_t - \mathrm{sg}\left[A\right]}_2^2}} \;,$
%     }
%     \label{eq: vector quantization losses}
% \end{align}
% \begin{align}
%     \mathcal{L}_t =&~
%     \underset{\text{encoder+decoder}}{\underbrace{\log p(x_t | q_{[1,t]})}} + \underset{\text{codebook}}{\underbrace{\text{MSE}\left(\mathrm{sg}\left[H\right], A\right)^2}} + \nonumber\\
%     & \underset{\text{encoder}}{\underbrace{\beta~\text{MSE}\left(H, \mathrm{sg}\left[A\right]\right)^2}} \;,
%     \label{eq: vector quantization losses}
% \end{align}
\edit{
\begin{align}
    \resizebox{.44\textwidth}{!}{%
    $\mathcal{L}_t = \underset{\text{encoder+decoder}}{\underbrace{\log p(x_t | q_{[1,t]})}} + \underset{\text{codebook}}{\underbrace{\text{MSE}\left(\mathrm{sg}\left[h_t\right], A\right)}} + \underset{\text{encoder}}{\underbrace{\alpha~\text{MSE}\left(h_t, \mathrm{sg}\left[A\right]\right)}} \;,$
    }
    \label{eq: vector quantization losses}
\end{align}}%
% \begin{align}
%     \mathcal{L}_t =&~
%     \underset{\text{encoder+decoder}}{\underbrace{\log p(x_t | q_{[1,t]})}} + \nonumber\\
%     & \underset{\text{codebook}}{\underbrace{\frac{1}{KD} \sum_{t=1}^{T}\sum_{k=1}^{K}\sum_{i=1}^{D} \left(\mathrm{sg}\left[h_{t,i}\right] - a_{k,i}\right)^2}} + \nonumber\\
%     & \underset{\text{encoder}}{\underbrace{\frac{\beta}{KD}\sum_{t=1}^{T}\sum_{k=1}^{K}{\sum_{i=1}^{D} \left(h_{t,i} - \mathrm{sg}\left[a_{k,i}\right]\right)^2}}} \;,
%     \label{eq: vector quantization losses}
% \end{align}
\noindent where \edit{$\log p(x_t|q_{[1,t]})$ is a reconstruction likelihood term usually using a categorical distribution,} $\mathrm{sg}[x] = x$ is the so-called stop-gradient operator \edit{which acts as the identity function during the forward pass but \ci{is assumed to have}
partial derivatives all equal to zero during the backward pass}, $\alpha$ is a scalar hyperparameter, 
% and we define $\text{MSE}(H,A) = \frac{1}{KD} \sum_{t=1}^{T}\sum_{k=1}^{K}\sum_{i=1}^{D}\left(h_{t,i} - a_{k,i}\right)^2$. 
\edit{and we define $\text{MSE}(h_t,A) = \frac{1}{KD} \sum_{k=1}^{K}\sum_{i=1}^{D}\left(h_{t,i} - a_{k,i}\right)^2$. 
The loss for a full sequence is the sum or mean over all $\mathcal{L}_t$.} 
% We note that the subtraction of a matrix from a vector in \cref{eq: vector
% quantization losses} implies the use of broadcasting.

These learned discrete representations have been shown to capture high-level
speech information closely related to phonemes, and are useful for
applications such as speaker conversion~\cite{chorowski2019unsupervised}.
Vector quantization is \edit{not} exclusive to VQ-VAE but has seen
widespread application within SSL for regularization purposes and to define
targets for the pretext task. We will cover these applications below.
%We highlight the models using vector-quantization in the column \textbc{{qtz}} in \cref{tab:model-taxonomy}.

\edit{The Gumbel softmax \cite{Gumbel-Softmax} is another frequently used approach for obtaining a discrete representation space, and has also been used for AEs \cite{eloff2019unsupervised}. In addition to the approaches discussed above, several other works on speech representation learning take inspiration from the AE framework \cite{zeiler2013rectified, badino2014auto, badino2015discovering, kamper2015unsupervised, renshaw2015comparison, settle2019_a2w}.}

\begin{comment}
\paragraph{Regularization}
In the generative-based approaches, some regularization in the encoder $f(\cdot)$ or decoder $g(\cdot)$ is required to encourage the model to utilize global speech information and prevent naive copying input when reconstruction. 
The models below have been enhanced by regularization methods. 
\begin{itemize}
%\item VAE, VQ-VAE
    \item DeCoAR 2.0~\cite{ling2020decoar} presents a deep contextualized acoustic representation learning approach with the addition of a vector quantization (VQ) layer.
    \item In VQ-APC~\cite{chung20e_interspeech}, a VQ layer is used with the APC objective, which imposes a bottleneck and forces the model to learn better representations.
    \item Two dropout regularization methods, attention dropout and layer dropout, are introduced to TERA~\cite{luo2021drop}. 
\end{itemize}
\end{comment}

@@

Autoregressive predictive coding (APC)~\cite{chung2019unsupervised,
chung2020generative} \edit{takes inspiration from the classic Linear Predictive Coding (LPC) approach for speech feature extraction~\cite{LPC} and autoregressive language models (LM) for text, where the model learns to predict future information from past}.
% This approach can be viewed as training a speech version of an LM.
%Just like an LM for text, this approach uses an encoder model to encode temporal information of past acoustic sequence, and the decoder then predicts future frames.
%In this approach, given $X=\{x_1,x_2, ..., x_T\}$, $\widetilde{X}$ is the acoustic features up to time step $t$, that is, $X_{[0,t]} = \{x_1,x_2, ..., x_{t}\}$.
\edit{A function} $f(\cdot)$ reads the input sequence $X_{[1,t]}$ and \edit{outputs a representation sequence} $H_{[1,t]}$.
The \edit{auxiliary module} $g(\cdot)$ is a linear projection layer which takes the last vector of $H_{[1,t]}$ as input to \edit{approximate} $x_{t+c}$, where $c \geq 1$. Thus, $c$ indicates how many timesteps the model predicts ahead. The \edit{modules} $f(\cdot)$
and $g(\cdot)$ are jointly learned to minimize {the \ensuremath{L_1} loss between $x_{t+c}$ and its approximation $\hat{x}_{t+c}$}. APC is formulated as 
%\begin{align}
    %Z_{[0,t-1]} &= f(X_{[0,t-1]}) \\
    %\hat{\mathbf{x}}_{t+c} &= g(\mathbf{z}_{t-1}) \label{eq:c} \\
    %\mathcal{L}_t &= \lVert \hat{\mathbf{x}}_{t+c} - \mathbf{x}_{t+c} \rVert_1\enspace.
%\end{align}
\begin{align}
    H_{[1,t]} &= f(X_{[1,t]}) , \\
    \hat{x}_{t+c} &= g(h_{t}) \label{eq:c} , \\
    \mathcal{L}_t &= \lVert \hat{x}_{t+c} - x_{t+c} \rVert_1 .
\end{align}
In text-based autoregressive LMs, $c$ is set to $1$ \edit{to enable autoregressive generation}. However, due to the smoothness of the speech signal, neighboring acoustic features are usually similar. Depending on the downstream task, we are often interested in learning so-called \emph{slow features} that typically span multiple input frames~\cite{wiskott2002slow}. Even the smallest linguistic units of speech---phonemes---span $0.07$~seconds on average in the English TIMIT dataset~\cite{garofolo_timit_1993},   whereas spectrogram frames $\mathbf{x}_t$ are typically computed at $0.01$ second intervals. Thus, simply predicting the next frame constitutes a trivial pretext task for APC; the original work finds that $c=3$ performs well. 
%\begin{equation}
  %  L = \sum_{X} \sum_{t} d(x_{t+c}, g(h_t)),
%   L = d(x_{t+c}, g( f(X_{[0,t-1]}) )),
%\end{equation}
%where $d(x_{t+c}, g(h_t))$ is the distance between $x_{t+c}$ and  $g(h_t)$. 
In \cite{chung2020improved}, the APC objective is extended to multi-target
training. The new objective generates both past and future frames conditioned
on previous context. 
In VQ-APC~\cite{chung20e_interspeech}, quantization is used with the APC
objective, which imposes an information bottleneck serving as a regularizer.

>>>>

\section{Speech Representation Learning Paradigms} \label{sec:approach}
Due to the characteristics of speech, SSL pretext tasks developed for CV
and NLP may not directly apply to speech.
Below we summarize the characteristics of speech as compared to CV and NLP.
\begin{itemize}
\item \textit{Speech is a sequence.} 
Unlike CV, in which an image usually has a fixed size representation, it is
natural to represent a speech utterance as a variable-length sequence. 
Therefore, pretext tasks developed for CV cannot generally be directly
applied to speech.

\item \textit{Speech is a long sequence without segment boundaries.} 
Both text and speech can be represented as sequences. From this viewpoint, it
is natural to apply learning approaches developed for text directly to
speech. 
In NLP, morpheme-like tokens are widely used as sequence units in
pre-training. The standard BERT takes 512 morpheme-like tokens as
input, usually covering a paragraph including several sentences. 
However, speech signals consist of sound pressure measurements with thousands
of samples per second, resulting in sequences much longer than those for text. Even
spectral representations which reduce the sequence length can have hundreds of
frames per second.
Processing such sequences with typical neural network architectures like
Transformers can result in problems with running time and memory requirements. 
One could gather consecutive frames to form shorter segments,
but unlike text, there is no obvious segmentation for unlabeled
speech.
\item \textit{Speech is continuous.} 
In NLP, it is common to use a pretext task that models a categorical
distribution of masked or future inputs. Since text is easily broken down into
individual tokens such as words, subwords, or characters, it is
straightforward to define a finite vocabulary for such tasks.
However, this idea does not apply to speech modeling because speech
signals are continuous; 
in this sense there is no such thing as a speech vocabulary. 
\item \textit{Speech processing tasks are diverse.}
Building generalizable self-supervised representation models for diverse speech
processing tasks is challenging.
Speech contains rich, hierarchical information, and different speech tasks
may require mutually orthogonal information.
For example, speech recognition requires a model that extracts content information
but ignores speaker information; in contrast, speaker recognition 
requires a model that extracts speaker information but removes content information.
Therefore, it is challenging to define a self-supervised model whose
representations are suitable for both speech recognition and speaker
recognition. Analogous considerations apply within CV and NLP.
\end{itemize}

In the sections below, we group modern SSL pretext tasks designed for speech
into three main categories: \textit{generative} approaches,
\textit{contrastive} approaches and \textit{predictive} approaches. 
\Cref{fig:timeline} shows a timeline of the models covered in these sections
with each model colored according to our categorization. 
\Cref{table:pretext} summarizes model pretext tasks along within the categories.

@@

We use $Q = \{q_1,q_2, ..., q_T\}$ to denote representations that are quantized via codebook learning. Alternatively, discrete representations may take the form of one-hot vectors, or the equivalent integer IDs, which we denote by $C = \{c_1,c_2, ..., c_T\}$. We use a circumflex to denote that, for instance, $\hat{x}_t$ is an approximation of $x_t$. Finally, we often use a subscript when defining a loss, $\mathcal{L}_i$, to imply that the total loss is computed as a sum over $i$, unless otherwise stated.

@@

\edit{After pre-training,} the representation model $f(\cdot)$ can be fine-tuned for a downstream task directly or used to extract features which are fed to another model, as visualized in \cref{fig:SSL_framework}. It is not uncommon to use the output representation~$H$, but often representations from hidden layers of $f(\cdot)$ are better suited \cite{pasad_layerwise_2021}.

\begin{figure*}[t!]
    \centering
    \includegraphics[width=0.98\textwidth]{paper_review/model_timeline.pdf}
   % \includegraphics[width=0.98\textwidth]{paper_review/model_timeline_wo_dccae.pdf}
	 \caption{A selection of models listed according to first publication date
	 on arXiv or conference submission date when this clearly precedes the
	 former. The models are categorized as generative, contrastive, or predictive.
	 In addition, some models are characterized as embedding models or
	 multi-modal models, although most learn frame-level
	 representations from speech only. Some models use a mixture of generative
	 and contrastive tasks. For instance, PASE and PASE+ use a multi-task setup,
	 but find that generative tasks are the most important for downstream
	 task performance~\cite{pascual_learning_2019}.}
    \label{fig:timeline}
\end{figure*}

@@

@@

In this category, the pretext task is to generate, or reconstruct, the input data based on some limited view. This includes predicting future inputs from past inputs, masked from unmasked, or the original from some other corrupted view.  ``Generative'' as used in this paper hence refers to models that target the original input in their pretext task. Note that this differs from generative models, which learn \ci{distributions that allow} to sample new data.

@@

\paragraph{Autoencoding} 
Since their introduction in the mid-1990s~\cite{hinton_autoencoders_1994}, autoencoders (AEs)
have played an essential role in learning distributed latent representations of
sensory data. 
As described above, AEs consist of an encoder and decoder; the pretext task
is to reconstruct the given input. The most common type of AE places an
information bottleneck on the latent representation by simply having fewer
hidden units available than input features. This forces the model to discard
low-level details and disccourages the learning of trivial solutions. Other models add
regularization to the latent space to further improve the quality of the
learned representations.
For instance, denoising autoencoders (DAEs) learn latent representations by
reconstructing from input corrupted by noise~\cite{vincent_stacked_2010}. 
\edit{The Variational Autoencoder (VAE) is a probabilistic version of the AE which defines the latent representation via a posterior distribution over stochastic latent variables \cite{kingma_autoencoding_2014,rezende_stochastic_2014}. VAEs have been applied to speech in numerous works \cite{chung_recurrent_2015, fraccaro_sequential_2016, hsu_learning_2017, hsu_unsupervised_2017, aksan_stcn_2019}}.
The vector-quantized variational autoencoder (VQ-VAE) is another model in this
category~\cite{oord_neural_2018};
it extends the original VAE~\cite{kingma_autoencoding_2014} with a novel
parameterization of the posterior distribution for discrete latent
representations. 
\edit{The VQ-VAE has been instrumental in generative speech modelling and recent work on generative spoken language modeling has successfully combined the idea of a discrete latent space with self-supervised learning \cite{polyak_speech_2021, kharitonov_textfree_2021, nguyen_generative_2022}.}

Specifically, in the VQ-VAE, the continuous representation vector $h_t$ at the output of the encoder is quantized by mapping it to a codebook vector, which is then used as the input to the decoder. This operation is non-differentiable and the \edit{gradients of the loss with respect to the encoder parameters} must be obtained by approximation. In the VQ-VAE this is done using the straight-through estimator~\cite{bengio_estimating_2013}, i.e., the gradients \edit{with respect to the} encoder output are taken to be equal to those \edit{with respect to the} decoder input \ci{(i.e., the quatization step is ignored)}. Given a learned codebook $A\in\mathbb{R}^{K \times D}$, where $K$ is the codebook size and $D$ is the dimensionality of each codebook vector $a_k$, the quantized representation $q_t$ of $h_t$ is obtained as
\begin{align}
    q_t = a_k, \text{ where } k=\arg\min_j \norm{h_t - a_j}_2 .
\end{align}
\edit{The decoder $g(\cdot)$ is an autoregressive model that takes $q_{[1,t]}$ as input to generate $x_t$ \cite{oord_wavenet_2016}}. 
Codebook learning is facilitated by a two-term auxiliary loss similar to
classical vector quantization dictionary 
learning~\cite{burton_generalization_1983, soong_vector_1985}. 
Gradients for the codebook vectors are given solely by a term that moves
codebook vectors $a_k$ closer to the non-quantized vectors $h_t$. A so-called
\emph{commitment term} is added to ensure that non-quantized vectors do not grow
unboundedly by enforcing the encoder to keep them close to a codebook vector.
This commitment term is optimized only by the encoder. The
total VQ-VAE loss for a single timestep is

\begin{align}
    \mathcal{L}_t = \underset{\text{encoder+decoder}}{\underbrace{\log p(x_t | q_{[1,t]})}} + \underset{\text{codebook}}{\underbrace{\text{MSE}\left(\mathrm{sg}\left[h_t\right], A\right)}} + \underset{\text{encoder}}{\underbrace{\alpha~\text{MSE}\left(h_t, \mathrm{sg}\left[A\right]\right)}}
    \label{eq: vector quantization losses}
\end{align}%
\noindent where \edit{$\log p(x_t|q_{[1,t]})$ is a reconstruction likelihood term usually using a categorical distribution,} $\mathrm{sg}[x] = x$ is the so-called stop-gradient operator \edit{which acts as the identity function during the forward pass but \ci{is assumed to have}
partial derivatives all equal to zero during the backward pass}, $\alpha$ is a scalar hyperparameter, 
\edit{and we define $\text{MSE}(h_t,A) = \frac{1}{KD} \sum_{k=1}^{K}\sum_{i=1}^{D}\left(h_{t,i} - a_{k,i}\right)^2$. 
The loss for a full sequence is the sum or mean over all $\mathcal{L}_t$.}

These learned discrete representations have been shown to capture high-level
speech information closely related to phonemes, and are useful for
applications such as speaker conversion~\cite{chorowski_unsupervised_2019}.
Vector quantization is \edit{not} exclusive to VQ-VAE but has seen
widespread application within SSL for regularization purposes and to define
targets for the pretext task. We will cover these applications below.

\edit{The Gumbel softmax \cite{jang_categorical_2017} is another frequently used approach for obtaining a discrete representation space, and has also been used for AEs \cite{eloff_unsupervised_2019}. In addition to the approaches discussed above, several other works on speech representation learning take inspiration from the AE framework \cite{zeiler_rectified_2013, badino_autoencoder_2014, badino_discovering_2015, kamper_unsupervised_2015, renshaw_comparison_2015, settle_acoustically_2019}.}

\begin{comment}
\paragraph{Regularization}
In the generative-based approaches, some regularization in the encoder $f(\cdot)$ or decoder $g(\cdot)$ is required to encourage the model to utilize global speech information and prevent naive copying input when reconstruction. 
The models below have been enhanced by regularization methods. 
\begin{itemize}
%\item VAE, VQ-VAE
    \item DeCoAR 2.0~\cite{ling2020decoar} presents a deep contextualized acoustic representation learning approach with the addition of a vector quantization (VQ) layer.
    \item In VQ-APC~\cite{chung_vector-quantized_2020}, a VQ layer is used with the APC objective, which imposes a bottleneck and forces the model to learn better representations.
    \item Two dropout regularization methods, attention dropout and layer dropout, are introduced to TERA~\cite{luo_drop_2021}. 
\end{itemize}
\end{comment}

@@

Autoregressive predictive coding (APC)~\cite{chung_unsupervised_2019,
chung_generative_2020} \edit{takes inspiration from the classic Linear Predictive Coding (LPC) approach for speech feature extraction~\cite{oshaughnessy_linear_1988} and autoregressive language models (LM) for text, where the model learns to predict future information from past}.
\edit{A function} $f(\cdot)$ reads the input sequence $X_{[1,t]}$ and \edit{outputs a representation sequence} $H_{[1,t]}$.
The \edit{auxiliary module} $g(\cdot)$ is a linear projection layer which takes the last vector of $H_{[1,t]}$ as input to \edit{approximate} $x_{t+c}$, where $c \geq 1$. Thus, $c$ indicates how many timesteps the model predicts ahead. The \edit{modules} $f(\cdot)$
and $g(\cdot)$ are jointly learned to minimize {the \ensuremath{L_1} loss between $x_{t+c}$ and its approximation $\hat{x}_{t+c}$}. APC is formulated as 
\begin{align}
    H_{[1,t]} &= f(X_{[1,t]}) , \\
    \hat{x}_{t+c} &= g(h_{t}) \label{eq:c} , \\
    \mathcal{L}_t &= \lVert \hat{x}_{t+c} - x_{t+c} \rVert_1 .
\end{align}
In text-based autoregressive LMs, $c$ is set to $1$ \edit{to enable autoregressive generation}. However, due to the smoothness of the speech signal, neighboring acoustic features are usually similar. Depending on the downstream task, we are often interested in learning so-called \emph{slow features} that typically span multiple input frames~\cite{wiskott_slow_2002}. Even the smallest linguistic units of speech---phonemes---span $0.07$~seconds on average in the English TIMIT dataset~\cite{garofolo_timit_1993},   whereas spectrogram frames $\mathbf{x}_t$ are typically computed at $0.01$ second intervals. Thus, simply predicting the next frame constitutes a trivial pretext task for APC; the original work finds that $c=3$ performs well. 
In \cite{chung_improved_2020}, the APC objective is extended to multi-target
training. The new objective generates both past and future frames conditioned
on previous context. 
In VQ-APC~\cite{chung_vector-quantized_2020}, quantization is used with the APC
objective, which imposes an information bottleneck serving as a regularizer.

A drawback of APC is that it encodes information only from previous timesteps
and not the entire input.
DeCoAR~\cite{ling_deep_2020} combines the bidirectionality of the popular NLP model
ELMo~\cite{peters_deep_2018} and the reconstruction objective of APC to alleviate
this issue and allow encoding information from the entire input. 
It uses a forward LSTM $f_1(\cdot)$ to encode $X_{[1,t]}$ and a backward LSTM
$f_2(\cdot)$ to encode $X_{[t+k,T]}$, where $k>1$: 
\begin{gather}
    H_{[1,t]} = f_1(X_{[1,t]}), \\
    H^\prime_{[t+k,T]} = f_2(X_{[t+k,T]}), \\
    \hat{X}_{[t+1,t+k-1]} = g(h_t, h^\prime_{t+k}).
\end{gather}
The input feature vector used in the downstream tasks is the concatenation of
$h_{t}$ and $h^\prime_{t}$.

\paragraph{Masked Reconstruction}

Masked reconstruction is largely inspired by the masked language model (MLM) task from BERT~\cite{devlin_bert:_2018}. During BERT pre-training, some tokens in the input sentences are masked by randomly replacing them by a learned masking token or another input token. The model learns to reconstruct the masked tokens from the non-masked tokens. Recent work has explored similar pretext tasks \edit{for speech representation learning}. Similar to the DeCoAR model described above, this allows a model to learn contextualized representations that encode information from the entire input. While we here focus on the models that reconstruct the masked input, it is important to note that masking has also been used extensively for contrastive (\cref{contrastive_approaches}) and predictive (\cref{predictive_approaches}) models.
